Component response rate variation drives stability in large complex systems

### 2018-07-29 notebook/SI.Rmd
Discussion draft written and SI edits

Overall, the ratio of stable `A1` matrices to stable `A0` matrices found is greater than 1 (compare column 5 to column 3), and this ratio increases with increasing $S$ (column 1). Hence, more randomly created complex systems ($\mathbf{M}$) are stable given variation in $\gamma$ than when $\gamma = 1$. Note that feasibility results were ommited for the table above, but are [reported below](#Feasibility).

@@

```{r, echo = FALSE}
get_top_evo_out <- function(evo_out, size){
    highest <- max(evo_out);
    gammas  <- NULL;
    while(size > 0){
        pos <- which(evo_out == highest);
        len <- length(pos);
        nli <- NULL;
        for(i in 1:len){
            start          <- pos[i] + 1;
            end            <- pos[i] + highest;
            gammas[[size]] <- evo_out[start:end];
            size           <- size - 1;
            if(size == 0){
                break;
            }
        }
        highest <- highest - 1;
    }
    return(gammas);
}
plot_evo_out <- function(evo_out){
    gammas <- get_top_evo_out(evo_out, size = 9);
    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xlim = c(0, 1.1));
    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 1.1));
    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
          text = expression(paste("Component response rate (",gamma,")")));
    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
}
evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
evo_res <- evo_res[,-1];
kable(evo_res);
```

@@

While the foundational work of May[@May1972] applies broadly to complex networks, much attention has been given specifically to ecological networks of interacting species. In these networks, the matrix $\mathbf{M}$ is interpreted as a community matrix and each row and column is interpreted as a single species. The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $M_{ij}$, meaning that $\mathbf{M}$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015]. While May's original work[@May1972] considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $\mathbf{M}$ are negative), mutualist networks (all off-diagonal elements of $\mathbf{M}$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa)[@Allesina2012; @Allesina2015]. In general, competitor and mutualist networks tend to be unstable, while predator-prey networks tend to be highly stabilising.

@@

```{r, echo = FALSE}
cd_tab <- cdat[,2:7];
kable(cd_tab);
```

@@

```{r, echo = FALSE}
md_tab <- mdat[,2:7];
kable(md_tab);
```

@@

```{r, echo = FALSE}
pd_tab <- pdat[,2:7];
kable(pd_tab);
```

@@

Sensitivity of inter-connectivity (C) values {#connectance}
================================================================================

@@

```{r, echo = FALSE}
C3dat   <- C3dat[,-1];
C3d_tab <- C3dat[,1:7];
kable(C3d_tab);
```

@@

```{r, echo = FALSE}
C5dat   <- C5dat[,-1];
C5d_tab <- C5dat[,1:7];
kable(C5d_tab);
```

@@

```{r, echo = FALSE}
C7dat   <- C7dat[,-1];
C7d_tab <- C7dat[,1:7];
kable(C7d_tab);
```

@@

```{r, echo = FALSE}
C9dat   <- C9dat[,-1];
C9d_tab <- C9dat[,1:7];
kable(C9d_tab);
```

>>>>

Overall, the ratio of stable `A1` matrices to stable `A0` matrices found is greater than 1 whenever $S > 10$ (compare column 5 to column 3), and this ratio increases with increasing $S$ (column 1). Hence, more randomly created complex systems ($\mathbf{M}$) are stable given variation in $\gamma$ than when $\gamma = 1$. Note that feasibility results were omitted for the table above, but are [reported below](#Feasibility).

@@

```{r, echo = FALSE}
get_top_evo_out <- function(evo_out, size){
    highest <- max(evo_out);
    gammas  <- NULL;
    while(size > 0){
        pos <- which(evo_out == highest);
        len <- length(pos);
        nli <- NULL;
        for(i in 1:len){
            start          <- pos[i] + 1;
            end            <- pos[i] + highest;
            gammas[[size]] <- evo_out[start:end];
            size           <- size - 1;
            if(size == 0){
                break;
            }
        }
        highest <- highest - 1;
    }
    return(gammas);
}
plot_evo_out <- function(evo_out){
    gammas <- get_top_evo_out(evo_out, size = 9);
    for(i in 1:9){
        gammas[[i]] <- 2 * gammas[[i]]; # Account for mistake in return
    }
    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xlim = c(0, 2.1));
    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 2.1));
    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
          text = expression(paste("Component response rate (",gamma,")")));
    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
}
evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
evo_res <- evo_res[,-1];
kable(evo_res);
```

@@

While the foundational work of May[@May1972] applies broadly to complex networks, much attention has been given specifically to ecological networks of interacting species. In these networks, the matrix $\mathbf{A}$ is interpreted as a community matrix and each row and column is interpreted as a single species. The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $A_{ij}$, meaning that $\mathbf{A}$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015]. While May's original work[@May1972] considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $\mathbf{A}$ are negative), mutualist networks (all off-diagonal elements of $\mathbf{A}$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa)[@Allesina2012; @Allesina2015]. In general, competitor and mutualist networks tend to be unstable, while predator-prey networks tend to be highly stabilising.

@@

```{r, echo = FALSE}
cd_tab <- cdat[,2:8];
kable(cd_tab);
```

@@

```{r, echo = FALSE}
md_tab <- mdat[,2:8];
kable(md_tab);
```

@@

```{r, echo = FALSE}
pd_tab <- pdat[,2:8];
kable(pd_tab);
```

@@

Sensitivity of connectance (C) values {#connectance}
================================================================================

@@

```{r, echo = FALSE}
C3dat   <- C3dat[,-1];
C3d_tab <- C3dat[,1:8];
kable(C3d_tab);
```

@@

```{r, echo = FALSE}
C5dat   <- C5dat[,-1];
C5d_tab <- C5dat[,1:8];
kable(C5d_tab);
```

@@

```{r, echo = FALSE}
C7dat   <- C7dat[,-1];
C7d_tab <- C7dat[,1:8];
kable(C7d_tab);
```

@@

```{r, echo = FALSE}
C9dat   <- C9dat[,-1];
C9d_tab <- C9dat[,1:8];
kable(C9d_tab);
```
